---
output:
  html_document: default
  pdf_document: default
---
title: "Homework - 2"
author: "Group 7"
date: "10/18/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importing the libraries

``` {r message = FALSE}
library(dplyr)
library(readxl)
library(psych)
```

# Problem 1a

``` {r message = FALSE}
# Import the NHL excel file as a dataframe
NHL_data <- data.frame(read_xlsx("NHL.xlsx", sheet = "Data"))
# Select the columns 13-26, the 1st column is the index column from excel
df <- NHL_data[, 13:26]
# Use Parallel Analysis Scree Plots to figure out the number of factors to extract
fa.parallel(df, fa = "pc", n.iter = 100, show.legend = FALSE)
```


#It is appropriate to use 1 factor 
# 1b
``` {r}
# Performing PCA with varimax orthogonal rotation
pc <- principal(df, nfactors = 1, rotate = "none", scores = TRUE)
pc
```

#Rotating the components
# Problem 1c
``` {r}
# PCA with varimax rotation
pc <- principal(df, nfactors = 1, rotate = "varimax", scores = TRUE)
pc
```

# Problem 1d
#Displaying the components score

``` {r}
head(pc$scores)
```


# Problem 1e
``` {r}
# Plotting the components and analyze them
factor.plot(pc, labels = colnames(df))
rm(list = ls())
```

# Problem 1f
#Interpretations -
# - The PCA loads all the components except face and PPP.

# - It is unclear as to what the component signifies without a deeper analysis

---

# Problem 2a

``` {r}
glass_data <- data.frame(read_xlsx("Glass Identification Data.xlsx",
  sheet = "Glass Data"
))
fa.parallel(glass_data[, 2:10], fa = "pc", n.iter = 100, show.legend = FALSE)
```


#The Skree Plot and the fa.parallel() function overlaps and shows us that the nfactors as 4.

# Problem 2b
``` {r}
# Perform PCA without rotation
pc <- principal(glass_data[, 2:10], nfactors = 4, rotate = "none", scores = TRUE)
pc
```



# Problem 2c
#Rotating the components

``` {r}
# Performing PCA with rotation
pc <- principal(glass_data[, 2:10], nfactors = 4, rotate = "varimax", scores = TRUE)
pc
```
# Problem 2d
``` {r}
head(pc$scores)
```

# Problem 2e
``` {r}
# Plotting the components to interpret the results
factor.plot(pc,
  choose = c(1),
  labels = colnames(glass_data[, 2:10]),
  title = "PCA Component 1"
)
factor.plot(pc,
  choose = c(2),
  labels = colnames(glass_data[, 2:10]),
  title = "PCA Component 2"
)
factor.plot(pc,
  choose = c(3),
  labels = colnames(glass_data[, 2:10]),
  title = "PCA Component 3"
)
factor.plot(pc,
  choose = c(4),
  labels = colnames(glass_data[, 2:10]),
  title = "PCA Component 4"
)
factor.plot(pc, choose = c(1, 2), labels = colnames(glass_data[, 2:10]))
factor.plot(pc, choose = c(1, 3), labels = colnames(glass_data[, 2:10]))
factor.plot(pc, choose = c(1, 4), labels = colnames(glass_data[, 2:10]))
factor.plot(pc, choose = c(2, 3), labels = colnames(glass_data[, 2:10]))
factor.plot(pc, choose = c(2, 4), labels = colnames(glass_data[, 2:10]))
factor.plot(pc, choose = c(3, 4), labels = colnames(glass_data[, 2:10]))
factor.plot(pc, labels = colnames(glass_data[, 2:10]))
rm(list = ls())
```

#Interpretations from the plot - 
#- PC1 - Indicates RI, CA, Mg, Al and K. It also signifies glass with high refractive index.
#- PC2 - Indicates Mg, Al and Ba. Signifies glass with high Magnesium, Aluminum and Barium concentration.
#- PC3 - Indicates Na, K, Ba and Fe. Signifies glass with high Sodium, Potassium, Barium and Iron concentration.
#- PC4 -  Indicates RI and Si. Signifies glass with high Refractive Index and Silicon concentration.

---

# Problem 3a

``` {r}
"Harman75.cor" <-
  structure(list(cov = structure(c(
    1, 0.318, 0.403, 0.468, 0.321,
    0.335, 0.304, 0.332, 0.326, 0.116, 0.308, 0.314, 0.489, 0.125,
    0.238, 0.414, 0.176, 0.368, 0.27, 0.365, 0.369, 0.413, 0.474,
    0.282, 0.318, 1, 0.317, 0.23, 0.285, 0.234, 0.157, 0.157, 0.195,
    0.057, 0.15, 0.145, 0.239, 0.103, 0.131, 0.272, 0.005, 0.255,
    0.112, 0.292, 0.306, 0.232, 0.348, 0.211, 0.403, 0.317, 1, 0.305,
    0.247, 0.268, 0.223, 0.382, 0.184, -0.075, 0.091, 0.14, 0.321,
    0.177, 0.065, 0.263, 0.177, 0.211, 0.312, 0.297, 0.165, 0.25,
    0.383, 0.203, 0.468, 0.23, 0.305, 1, 0.227, 0.327, 0.335, 0.391,
    0.325, 0.099, 0.11, 0.16, 0.327, 0.066, 0.127, 0.322, 0.187,
    0.251, 0.137, 0.339, 0.349, 0.38, 0.335, 0.248, 0.321, 0.285,
    0.247, 0.227, 1, 0.622, 0.656, 0.578, 0.723, 0.311, 0.344, 0.215,
    0.344, 0.28, 0.229, 0.187, 0.208, 0.263, 0.19, 0.398, 0.318,
    0.441, 0.435, 0.42, 0.335, 0.234, 0.268, 0.327, 0.622, 1, 0.722,
    0.527, 0.714, 0.203, 0.353, 0.095, 0.309, 0.292, 0.251, 0.291,
    0.273, 0.167, 0.251, 0.435, 0.263, 0.386, 0.431, 0.433, 0.304,
    0.157, 0.223, 0.335, 0.656, 0.722, 1, 0.619, 0.685, 0.246, 0.232,
    0.181, 0.345, 0.236, 0.172, 0.18, 0.228, 0.159, 0.226, 0.451,
    0.314, 0.396, 0.405, 0.437, 0.332, 0.157, 0.382, 0.391, 0.578,
    0.527, 0.619, 1, 0.532, 0.285, 0.3, 0.271, 0.395, 0.252, 0.175,
    0.296, 0.255, 0.25, 0.274, 0.427, 0.362, 0.357, 0.501, 0.388,
    0.326, 0.195, 0.184, 0.325, 0.723, 0.714, 0.685, 0.532, 1, 0.17,
    0.28, 0.113, 0.28, 0.26, 0.248, 0.242, 0.274, 0.208, 0.274, 0.446,
    0.266, 0.483, 0.504, 0.424, 0.116, 0.057, -0.075, 0.099, 0.311,
    0.203, 0.246, 0.285, 0.17, 1, 0.484, 0.585, 0.408, 0.172, 0.154,
    0.124, 0.289, 0.317, 0.19, 0.173, 0.405, 0.16, 0.262, 0.531,
    0.308, 0.15, 0.091, 0.11, 0.344, 0.353, 0.232, 0.3, 0.28, 0.484,
    1, 0.428, 0.535, 0.35, 0.24, 0.314, 0.362, 0.35, 0.29, 0.202,
    0.399, 0.304, 0.251, 0.412, 0.314, 0.145, 0.14, 0.16, 0.215,
    0.095, 0.181, 0.271, 0.113, 0.585, 0.428, 1, 0.512, 0.131, 0.173,
    0.119, 0.278, 0.349, 0.11, 0.246, 0.355, 0.193, 0.35, 0.414,
    0.489, 0.239, 0.321, 0.327, 0.344, 0.309, 0.345, 0.395, 0.28,
    0.408, 0.535, 0.512, 1, 0.195, 0.139, 0.281, 0.194, 0.323, 0.263,
    0.241, 0.425, 0.279, 0.382, 0.358, 0.125, 0.103, 0.177, 0.066,
    0.28, 0.292, 0.236, 0.252, 0.26, 0.172, 0.35, 0.131, 0.195, 1,
    0.37, 0.412, 0.341, 0.201, 0.206, 0.302, 0.183, 0.243, 0.242,
    0.304, 0.238, 0.131, 0.065, 0.127, 0.229, 0.251, 0.172, 0.175,
    0.248, 0.154, 0.24, 0.173, 0.139, 0.37, 1, 0.325, 0.345, 0.334,
    0.192, 0.272, 0.232, 0.246, 0.256, 0.165, 0.414, 0.272, 0.263,
    0.322, 0.187, 0.291, 0.18, 0.296, 0.242, 0.124, 0.314, 0.119,
    0.281, 0.412, 0.325, 1, 0.324, 0.344, 0.258, 0.388, 0.348, 0.283,
    0.36, 0.262, 0.176, 0.005, 0.177, 0.187, 0.208, 0.273, 0.228,
    0.255, 0.274, 0.289, 0.362, 0.278, 0.194, 0.341, 0.345, 0.324,
    1, 0.448, 0.324, 0.262, 0.173, 0.273, 0.287, 0.326, 0.368, 0.255,
    0.211, 0.251, 0.263, 0.167, 0.159, 0.25, 0.208, 0.317, 0.35,
    0.349, 0.323, 0.201, 0.334, 0.344, 0.448, 1, 0.358, 0.301, 0.357,
    0.317, 0.272, 0.405, 0.27, 0.112, 0.312, 0.137, 0.19, 0.251,
    0.226, 0.274, 0.274, 0.19, 0.29, 0.11, 0.263, 0.206, 0.192, 0.258,
    0.324, 0.358, 1, 0.167, 0.331, 0.342, 0.303, 0.374, 0.365, 0.292,
    0.297, 0.339, 0.398, 0.435, 0.451, 0.427, 0.446, 0.173, 0.202,
    0.246, 0.241, 0.302, 0.272, 0.388, 0.262, 0.301, 0.167, 1, 0.413,
    0.463, 0.509, 0.366, 0.369, 0.306, 0.165, 0.349, 0.318, 0.263,
    0.314, 0.362, 0.266, 0.405, 0.399, 0.355, 0.425, 0.183, 0.232,
    0.348, 0.173, 0.357, 0.331, 0.413, 1, 0.374, 0.451, 0.448, 0.413,
    0.232, 0.25, 0.38, 0.441, 0.386, 0.396, 0.357, 0.483, 0.16, 0.304,
    0.193, 0.279, 0.243, 0.246, 0.283, 0.273, 0.317, 0.342, 0.463,
    0.374, 1, 0.503, 0.375, 0.474, 0.348, 0.383, 0.335, 0.435, 0.431,
    0.405, 0.501, 0.504, 0.262, 0.251, 0.35, 0.382, 0.242, 0.256,
    0.36, 0.287, 0.272, 0.303, 0.509, 0.451, 0.503, 1, 0.434, 0.282,
    0.211, 0.203, 0.248, 0.42, 0.433, 0.437, 0.388, 0.424, 0.531,
    0.412, 0.414, 0.358, 0.304, 0.165, 0.262, 0.326, 0.405, 0.374,
    0.366, 0.448, 0.375, 0.434, 1
  ), .Dim = c(24, 24), .Dimnames = list(
    c(
      "VisualPerception", "Cubes", "PaperFormBoard", "Flags",
      "GeneralInformation", "PargraphComprehension", "SentenceCompletion",
      "WordClassification", "WordMeaning", "Addition", "Code",
      "CountingDots", "StraightCurvedCapitals", "WordRecognition",
      "NumberRecognition", "FigureRecognition", "ObjectNumber",
      "NumberFigure", "FigureWord", "Deduction", "NumericalPuzzles",
      "ProblemReasoning", "SeriesCompletion", "ArithmeticProblems"
    ), c(
      "VisualPerception", "Cubes", "PaperFormBoard", "Flags",
      "GeneralInformation", "PargraphComprehension", "SentenceCompletion",
      "WordClassification", "WordMeaning", "Addition", "Code",
      "CountingDots", "StraightCurvedCapitals", "WordRecognition",
      "NumberRecognition", "FigureRecognition", "ObjectNumber",
      "NumberFigure", "FigureWord", "Deduction", "NumericalPuzzles",
      "ProblemReasoning", "SeriesCompletion", "ArithmeticProblems"
    )
  )), center = c(
    0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
    0, 0, 0, 0, 0, 0, 0, 0, 0, 0
  ), n.obs = 145), .Names = c(
    "cov",
    "center", "n.obs"
  ))
fa.parallel(Harman74.cor$cov,
  n.obs = Harman74.cor$n.obs,
  n.iter = 100
)
```

#The Screeplot indicates nfactor of 4 


# Problem 3b
``` {r}
# Performing Factor Analysis
correlations <- cov2cor(Harman74.cor$cov)
fa <- fa(correlations, nfactors = 4, rotate = "none", scores = TRUE)
fa
```

# Problem 3c
#Rotating the factors

``` {r}
fa_orthogonal <- fa(correlations, nfactors = 4, rotate = "varimax", scores = TRUE)
fa_orthogonal
```

# Problem 3d
#Let's see the factor scores
``` {r}
fa_orthogonal$weights
```


#Plotting an orthogonal solution
# Problem 3e
``` {r}
factor.plot(fa_orthogonal, choose = c(1), labels = rownames(fa$loadings))
factor.plot(fa_orthogonal, choose = c(2), labels = rownames(fa$loadings))
factor.plot(fa_orthogonal, choose = c(3), labels = rownames(fa$loadings))
factor.plot(fa_orthogonal, choose = c(4), labels = rownames(fa$loadings))
factor.plot(fa_orthogonal)
```

Now let's plot an oblique solution
``` {r}
fa_oblique <- fa(Harman74.cor$cov, nfactors = 4, rotate = "promax", scores = TRUE)
fa.diagram(fa_oblique)
rm(list = ls())
```
# Problem 3e
#We can make the following observations - 

#- Factor 1 indicates that it is related to language related attributes like Sentence Completion, WordMeaning, Word Classification etc.

#- Factor 2 looks to be related to more general problem solving attributes like Deduction, Problem Reasoning, Cubes, Flage etc.

#- Factor 3 shows that it is related to mathematical attributes like Code, Counting Data, Addition, Number Recognition etc.

#- Factor 4 indicates that it is related to Word and Number recognition.

---

# Problem 4a

``` {r}
breast_cancer <- read_xlsx("breast-cancer-wisconsin.xlsx",
  sheet = "breast-cancer-wisconsin.csv"
)
breast_cancer <- select(breast_cancer, c(1, 2, 3, 4, 5, 6, 8, 9, 10, 11))
bc.cor <- cor(breast_cancer[, 2:9])
```

# Problem 4b
#Visualizing the correlation
``` {r}
pairs.panels(bc.cor)
```

#Interpretations from the plot -

# - The following variable pairs did not show any coorelation:
#  1. CT and NN
#  2. CT and MA
#  3. CT and SECS
#  4. CT and BC

#Analyzing the scree plot
``` {r}
fa.parallel(bc.cor, n.obs = 699)
```

# Problem 4c
# From the scree plot the number of factors is 1

``` {r}
fa <- fa(bc.cor, nfactors = 1, rotate = "none", n.obs = 699, fm = "pa")
fa
```

# Problem 4d
# Rotate the factors now 

``` {r}
fa_orthogonal <- fa(bc.cor, nfactors = 1, rotate = "varimax", fm = "pa", scores = TRUE)
fa_orthogonal
```
# Problem 4e
#Displaying the factors :
```{r}
fa_orthogonal$weights
```

# Problem 4f
# Plotting the orthogonal solution
``` {r}
factor.plot(fa_orthogonal, labels = colnames(bc.cor))
```

# Plotting the oblique solution

``` {r}
fa_oblique <- fa(bc.cor, nfactors = 1, rotate = "promax", fm = "pa", scores = TRUE)
fa_oblique
```

``` {r}
factor.plot(fa_oblique, labels = colnames(bc.cor))
fa.diagram(fa_oblique)
rm(list = ls())
```

# Interpretations from the plots - 

#The high loading scores and the test of hypothesis of the
#factor analysis indicate that one factor is sufficient to explain all the variables owing to their 
#high loading scores, indicating high values for communality.
#both the orthoganal and oblique plot reinforce the same.

---

# Problem 5a

``` {r}
vertebral_column_data <- data.frame(read_xlsx("Vertebral Column Data.xlsx",
  sheet = "column_3C"
))
fa.parallel(vertebral_column_data[, 1:6], n.iter = 100)
```
#Scree Plots denotes factor 3

# Problem 5b
``` {r}
# Calculate the distance matrix
d <- dist(vertebral_column_data[, 1:6])
# Perform Multidimensional Scaling
fit <- cmdscale(d, k = 3, eig = TRUE)
```

#The MDS has reduced the data to 3 dims

``` {r}
head(fit$points)
```

#This is the same as calculating 3 factors.


# Problem 5c
#Plotting the orthogonal solution 

``` {r}
pc <- principal(vertebral_column_data[, 1:6], nfactors = 3, rotate = "varimax")
factor.plot(pc, choose = c(1), labels = c(
  "pelvic incidence", "pelvic tilt",
  "lumbar lordosis angle", "sacral slope",
  "pelvic radius", " grade of spondylolisthesis."
))
factor.plot(pc, choose = c(2), labels = c(
  "pelvic incidence", "pelvic tilt",
  "lumbar lordosis angle", "sacral slope",
  "pelvic radius", " grade of spondylolisthesis."
))
factor.plot(pc, choose = c(3), labels = c(
  "pelvic incidence", "pelvic tilt",
  "lumbar lordosis angle", "sacral slope",
  "pelvic radius", " grade of spondylolisthesis."
))
factor.plot(pc, labels = c(
  "pelvic incidence", "pelvic tilt",
  "lumbar lordosis angle", "sacral slope",
  "pelvic radius", " grade of spondylolisthesis."
))
rm(list = ls())
```

#Interpretations-

#Component 1 loads pelvic incidence, lumbar lordosis angle, sacral slope and grade of spondylolisthesis. 

#Component 2 loads pelvic incidence and pelvic tilt.

#Component 3 loads sacral slope and pelvic radius.
